### 2018-03-27 Paper/ros.tex

\section{Reduced Order Surrogate}
\label{sec:ros}