### 2018-03-28 Paper/setup.tex
set up

\section{NEMD Set-up}
\label{sec:setup}

Non-equilibrium molecular dynamics (NEMD) simulations are performed in LAMMPS~\cite{Plimpton:2007}
by applying a thermal gradient by means of thermostats located at $L$/4 and $3L$/4 in a silicon bar of length, $L$. 
The atomic configuration as well as a schematic of the set-up is provided in Figure~\ref{fig:setup}.