### 2018-04-09 Paper/ack.tex
...

\section*{Acknowledgment}
\label{sec:ack}

M. Vohra and S. Mahadevan gratefully acknowledge funding support from the
National Science Foundation Award $\#$. Molecular
dynamics simulations were performed using resources at the advanced computing
center for research and education (ACCRE) at Vanderbilt University. 
M. Vohra is grateful to the ACCRE staff for their technical guidance
and support. He would also like to sincerely thank Dr. Alen Alexanderian for
insightful discussions pertaining to the derivative-based sensitivity analysis
aspect of this work.