### 2018-04-10 Slides/r1/md_si.tex
..

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  This Beamer template was created by Cameron Bracken.
%%  Anyone can freely use or modify it for any purpose
%%  without attribution.
%%
%%  Last Modified: January 9, 2009
%%

\documentclass[xcolor={x11names,table},compress,svgnames,mathserif]{beamer}

%% General document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{decorations.fractals}
%\usepackage{lmodern}
\usepackage{animate}
\usepackage{movie15}
\usepackage{bm}
\usepackage{pifont}
\usepackage{empheq}
\usepackage[many]{tcolorbox}
\usepackage{smartdiagram}
\usepackage[customcolors]{hf-tikz}
\usepackage{dashrule} % for dotted vertical line
%\usepackage{flexisym}

\usepackage[absolute,overlay]{textpos}

\pgfdeclarehorizontalshading{section shading}{2cm}{
color(0cm)=(LightSlateGrey);
color(2cm)=(gray!7);
color(3cm)=(LightSlateGrey!15)
}

% Custom block environment
% Custom block environment
\newenvironment<>{varblock}[2][.9\textwidth]{%
  \setlength{\textwidth}{#1}
  \begin{actionenv}#3%
    \def\insertblocktitle{#2}%
    \par%
    \usebeamertemplate{block begin}}
  {\par%
    \usebeamertemplate{block end}%
  \end{actionenv}}

% boxed equaton
\usepackage[many]{tcolorbox}
%\tcbuselibrary{skins}

\tcbset{highlight math style={enhanced,
  colframe=red!60!black,colback=yellow!50!white,arc=4pt,boxrule=1pt,
  }}

\newtcbox{\mybox}[1][]{nobeforeafter,math upper,tcbox raise base,
  enhanced,frame hidden,boxrule=0pt,interior style={top color=green!10!white,
  bottom color=green!10!white,middle color=green!50!yellow},
  fuzzy halo=1pt with green,drop large lifted shadow,#1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning,decorations.pathreplacing}
% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=purple!20, 
    text width=5.0em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw=none, fill=blue!20, anchor=north, 
    text width=9.0em, text centered]
    \tikzstyle{blockr} = [rectangle, draw, fill=blue!20, 
    text width=9.0em, text centered, rounded corners]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw=none, ellipse,fill=purple!20, node distance=3cm,
    minimum height=2em]

%% Beamer Layout %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\useoutertheme[subsection=false,shadow]{miniframes}
\useinnertheme{default}
%\usefonttheme{serif}
\usepackage{palatino}
\usepackage{xcolor}
\usepackage{amsmath}
\newcommand{\angstrom}{\textup{\AA}}

\setbeamerfont{title like}{shape=\scshape}
\setbeamerfont{frametitle}{shape=\scshape}
\setbeamertemplate{itemize items}[triangle] % if you wnat a circle
\setbeamertemplate{itemize subitem}[triangle]
\setbeamertemplate{navigation symbols}{}
%\setbeamertemplate{footline}[frame number]

\setbeamercolor{footlinecolor}{fg=white,bg=DeepSkyBlue4}
\defbeamertemplate*{footline}{infolines theme}
{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{footlinecolor}
 % {author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshortauthor
   % \usebeamerfont{author in head/foot}\insertshortauthor~~(\insertshortinstitute)
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,right]{footlinecolor}
  %{date in head/foot}%
   % \usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}
    \usebeamerfont{date in head/foot}{manav.vohra@vanderbilt.edu}\hspace*{2em}
    \insertframenumber{} / \inserttotalframenumber\hspace*{2ex} 
  \end{beamercolorbox}}%
  \vskip0pt%
}
%\setbeamercolor{section in head/foot}{fg=white, bg=DeepSkyBlue4}

\setbeamercolor*{lower separation line head}{bg=DeepSkyBlue4} 
\setbeamercolor*{normal text}{fg=black,bg=white} 
\setbeamercolor*{alerted text}{fg=red} 
\setbeamercolor*{example text}{fg=black} 
\setbeamercolor*{structure}{fg=black}

\setbeamercolor*{palette tertiary}{fg=black,bg=black!10} 
\setbeamercolor*{palette quaternary}{fg=black,bg=black!10}

\renewcommand{\(}{\begin{columns}}
\renewcommand*\footnoterule{}
\renewcommand{\)}{\end{columns}}
\newcommand{\<}[1]{\begin{column}{#1}}
\renewcommand{\>}{\end{column}}

\newcommand*\subitem{%
  \item[\color{DeepSkyBlue4}\scalebox{0.6}{\ding{228}}]}

  \newcommand*\subitemtwo{%
  \item[\color{LightSlateGrey!15}\scalebox{0.6}{\ding{228}}]}

\newcommand*\myitem{%
  \item[\color{DeepSkyBlue4}\scalebox{0.6}{\ding{110}}]}

  \newcommand*\myitemtwo{%
  \item[\color{LightSlateGrey!15}\scalebox{0.6}{\ding{110}}]}

\newcommand*\Myitem{%
  \item[\color{DeepSkyBlue4}\scalebox{0.9}{\ding{42}}]}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\p}{\partial}
\def\ol{\overline}
\def\no{\noindent}
\def\Vb{{\cal V}}
\def\Qd{\dot{Q}}
%\DeclareMathSymbol{\ast}{\mathbin}{symbols}{"03}
 \newcommand{\argmax}{\operatornamewithlimits{arg\,max}}

%-----------Parameter screening tikz styles -----------------------
   \tikzstyle{A} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=0.5cm,text centered, draw=black,
     fill=blue!30]
      \tikzstyle{B} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=0.5cm,text centered, draw=black,
     fill=red!30]
      \tikzstyle{p} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=0.5cm,text centered, draw=black,
     fill=yellow!30]
      \tikzstyle{q} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=0.5cm,text centered, draw=black,
     fill=green!30]
      \tikzstyle{alpha} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=0.5cm,text centered, draw=black,
     fill=magenta!30]
      \tikzstyle{lambda} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=0.5cm,text centered, draw=black,
     fill=cyan!30]
      \tikzstyle{gamma} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=0.5cm,text centered, draw=black,
     fill=orange!30]
 \tikzstyle{outer} = [rectangle, minimum width=1.5cm, minimum height=0.5cm,text centered, draw=black,dashed]    
 \tikzstyle{pool} = [rectangle, minimum width=2.0cm, minimum height=1.6cm,line width=1pt,
text centered, draw=black,dashed]

%% Algorithm %%%%%
\usepackage[linesnumbered]{algorithm2e}
\newcommand{\nosemic}{\renewcommand{\@endalgocfline}{\relax}}% Drop semi-colon ;
\newcommand{\dosemic}{\renewcommand{\@endalgocfline}{\algocf@endline}}% Reinstate semi-colon ;
\newcommand{\pushline}{\Indp}% Indent
\newcommand{\popline}{\Indm\dosemic}% Undent
\let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}% Remove line number for one
\SetKwRepeat{Do}{do}{while}
%%%%%%%%%%%%%%%%%

%---------------------QUOTATION--------------------------
\usepackage{etoolbox}
%\usepackage[svgnames]{xcolor}
\usepackage{framed}

% conditional for xetex or luatex
\newif\ifxetexorluatex
\ifxetex
  \xetexorluatextrue
\else
  \ifluatex
    \xetexorluatextrue
  \else
    \xetexorluatexfalse
  \fi
\fi
%
\ifxetexorluatex%
  \usepackage{fontspec}
  \usepackage{libertine} % or use \setmainfont to choose any font on your system
  \newfontfamily\quotefont[Ligatures=TeX]{Linux Libertine O} % selects Libertine as the quote font
\else
  \usepackage[utf8]{inputenc}
  \usepackage[T1]{fontenc}
  %\usepackage{libertine} % or any other font package
  \newcommand*\quotefont{\fontfamily{LinuxLibertineT-LF}} % selects Libertine as the quote font
\fi

\newcommand*\quotesize{30} % if quote size changes, need a way to make shifts relative
% Make commands for the quotes
\newcommand*{\openquote}
   {\tikz[remember picture,overlay,xshift=-3ex,yshift=-0.5ex]
   \node (OQ) {\quotefont\fontsize{\quotesize}{\quotesize}\selectfont``};\kern0pt}

\newcommand*{\closequote}[1]
  {\tikz[remember picture,overlay,xshift=-15ex,yshift={#1}]
   \node (CQ) {\quotefont\fontsize{\quotesize}{\quotesize}\selectfont''};}

% select a colour for the shading
\colorlet{shadecolor}{Azure}

\newcommand*\shadedauthorformat{\emph} % define format for the author argument

% Now a command to allow left, right and centre alignment of the author
\newcommand*\authoralign[1]{%
  \if#1l
    \def\authorfill{}\def\quotefill{\hfill}
  \else
    \if#1r
      \def\authorfill{\hfill}\def\quotefill{}
    \else
      \if#1c
        \gdef\authorfill{\hfill}\def\quotefill{\hfill}
      \else\typeout{Invalid option}
      \fi
    \fi
  \fi}
% wrap everything in its own environment which takes one argument (author) and one optional argument
% specifying the alignment [l, r or c]
%
\newenvironment{shadequote}[2][l]%
{\authoralign{#1}
\ifblank{#2}
   {\def\shadequoteauthor{}\def\yshift{-2ex}\def\quotefill{\hfill}}
   {\def\shadequoteauthor{\par\authorfill\shadedauthorformat{#2}}\def\yshift{2ex}}
\begin{snugshade}\begin{quote}\openquote}
{\shadequoteauthor\quotefill\closequote{\yshift}\end{quote}\end{snugshade}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[UQ: Nano-scale Phonon Heat Transfer]{\textbf{Characterizing Errors and Uncertainties in
NEMD Simulations of Phonon Transport}}
\thispagestyle{empty}
\pgfsetfillopacity{0.9}
\setbeamercolor{title}{bg=DeepSkyBlue4,fg=white}

%\subtitle{SUBTITLE}
\author[M. Vohra and S. Mahadevan]{Manav Vohra$^{\dag}$, Sankaran Mahadevan$^{\dag}$ \\ \vspace{2mm}
\scriptsize{Collaborators: Seungha Shin$^{\S}$, Ali Yousefzadi Nobakht$^{\S}$}}
\vspace{-1mm}
\institute{$^{\dag}$Vanderbilt University\\ \vspace{1mm}
$^{\S}$University of Tennessee, Knoxville}

\date{\today}

\begin{document}

%___________________________NEW SLIDE______________________________________
{
\setbeamertemplate{headline}{}

\begin{frame}[noframenumbering]

\titlepage
\vspace{-21mm}
\centering
%\scriptsize Institute for Computational Engineering and Sciences\\ \vspace{1mm}
%The University of Texas at Austin \\ \vspace{6mm}

%\footnotesize Seminar at SCI, University of Utah \\ \vspace{2mm}

\end{frame}
}

%___________________________NEW SLIDE______________________________________

\end{document}