### 2018-07-02 Paper/Rebuttal/rebuttal.tex
..

\documentclass[12pt]{article}
\input{def}
\usepackage{amsmath} % or simply amstext
\usepackage{amssymb}
\usepackage{enumitem}

%\usepackage[left=0.75in ,top=0.5in, bottom=0.5in, right=0.75in, footskip=0.5cm]{geometry}
\begin{document}

\parskip=6pt

\begin{center}
{\bf Summary of Modifications to HMT\_2018\_2060}\\[6pt]
{\bf Uncertainty Quantification in Non-Equilibrium Molecular Dynamics Simulations of Thermal Transport}\\[6pt]
By \\
Manav Vohra, Ali Yousefzadi Nobakht, Seungha Shin, Sankaran Mahadevan 
\end{center}

\baselineskip=22pt

\vspace*{1in}

We thank the reviewers for their assessment of our manuscript.
Below is a summary of modifications made to our original submission in
response to the reviewers suggestions.

Where possible, key modifications have been highlighted in blue in the revised
manuscript.

We hope that with these modifications
the paper is found suitable for publication in the {\it International Journal of Heat and Mass Transfer}.

\clearpage

\section*{Referee 1}
\begin{enumerate}[leftmargin=*,itemsep=24pt]
\item \textbf{Comment 1}

\end{enumerate}

%\bibliographystyle{unsrt}
%\bibliography{REFER}

\end{document}